---
title: "203A Final Project"
author: "Cyrene Arputhasamy"
date: "2022-12-05"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r}

```

```{r}
library(epiDisplay)
library(lattice)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
nhds <- read.csv("finaldata.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)
nhds.working <- data.frame(nhds)
```
```{r}
nhds.working %>%  # replace to your needs
  summarise_all(funs(sum(is.na(.))))
```
The only variable that contains NAs is the secondary expected source of payment. Not all individuals are expected to have a secondary source of payment, so this is reasonable. We aren't intending to use that variable so we can leave it as is. 
```{r}
nhds.working$OWNER_fac <- factor(nhds$OWNER, 
                     levels = c(1,2,3),
                     labels = c("Proprietary", "Government", "Nonprofit"))
tab1(nhds.working$OWNER_fac, sort.group = "increasing", cum.percent = TRUE, main = "Distribution of Hospital Ownership", horiz = TRUE, bar.values = "percent")
```
More than 80% of the data is from Nonprofit hospitals, including churches. 
```{r}
nhds.working$ESOP1_fac <- factor(nhds.working$ESOP1,
levels = c(1,2,3,4,5,6,7,8,9,10,99),
            labels = c("Worker's comp", "Medicare", "Medicaid", "Other government", "Blue Cross/Blue Shield", "HMO/PPO", "Other Private", "Self-pay", "No charge", "Other", "Not stated"))
tab1(nhds.working$ESOP1_fac, sort.group = "decreasing", cum.percent = TRUE, bar.values = "percent", main = "Distribution of Principal Expected Source of Payment", cex.main = .75, cex.axis = .75, cex.name = .75, cex.lab = .75)
```
```{r}
nhds.working$YEAR_fac <- factor(nhds.working$YEAR,
levels = c(0,1,2,3,4,5,6,7),
            labels = c("2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007"))
df1<- nhds.working %>%
  group_by(YEAR_fac, OWNER_fac, ESOP1_fac) %>%
  summarize(Freq=n())
p <- ggplot(df1, aes(fill=OWNER_fac, y=Freq, x=YEAR_fac)) + 
    geom_bar(position="fill", stat="identity")
p2 <- ggplot(df1, aes(fill=ESOP1_fac, y=Freq, x=YEAR_fac)) + 
    geom_bar(position="fill", stat="identity")
p + labs(title = "Distribution Between 2000 to 2007", subtitle = "Hospital Ownership", x = "Years", y = "Frequency") + scale_fill_discrete(name = "Hospital Types")
p2 + labs(subtitle = "Principal Expected Source of Payment", x = "Years", y = "Frequency") + scale_fill_discrete(name = "Payment Types")
```

```{r}
table(nhds.working$OWNER_fac, nhds.working$ESOP1_fac )
chisq.test(nhds.working$OWNER_fac, nhds.working$ESOP1_fac)
```
By this test we can conclude that hospital ownership and principal expected source of income are not independently related. For example, the lowest frequency observed in Proprietary hospitals (hospitals which are for-profit) see individuals who are expected to have no charge. 
```{r}
heart <- filter(nhds.working, dx1_pre == 428)
tab1(heart$OWNER_fac, sort.group = "increasing", cum.percent = TRUE, main = "Distribution of Hospital Ownership in Patients Diagnosed with Heart Failure", horiz = TRUE, bar.values = "percent", cex.main = .75, cex.axis = .75, cex.name = .75, cex.lab = .75)
tab1(heart$ESOP1_fac, sort.group = "decreasing", cum.percent = TRUE, bar.values = "percent", main = "Distribution of Principal Expected Source of Payment in Patients Diagnosed with Heart Failure", cex.main = .55, cex.axis = .75, cex.name = .75, cex.lab = .75)
```
Distributions of hospital ownership are very similar to the distributions found in the entire dataset. Distributions of principal expected source of payment are skewed further toward Medicare in those with heart failure than in the entire dataset.
```{r}
heart2 <- heart[!apply(heart == "", 1, all),]
sort(table(heart2$PD1), decreasing = TRUE)[1:6]
```
The top five procedure codes in those with heart failure are as follows:
Code 8872: Diagnostic Ultrasound of the Heart
Code 3995: Hemodialysis
Code 3722: Left Heart Cardiac Catheterization
Code 9904: Transfusion of Packed Cells
Code 3491: Thoracentesis 
Only code 8872 and 3722 have to do with the heart so we'll continue with those codes.
```{r}
heart3 <- filter(heart2, PD1 == 8872 | PD1 == 3722)
m_heart <- lm(PD1 ~ OWNER_fac + ESOP1_fac, data = heart3)
summary(m_heart)
```
According to the linear model summary, each level of hospital ownership is a statistically significant predictor of either of the two chosen procedures in patients diagnosed with heart failure. Some levels of principal payment method are not significant like Blue Cross/Blue Shield, HMO/PPO and other private means of payment. It is seen that whether the patient is paying by government funds (Medicare, Medicaid, other government) are statically significant predictors of these procedures. 

To assess whether the linear model is reliable, we need to run some model diagnostics and check for (1) linearity, (2) nearly normal residuals, and (3) constant variability.
```{r}
#Linearity
ggplot(data = m_heart, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")

#Nearly Normal Residuals
ggplot(data = m_heart, aes(x = .resid)) +
  geom_histogram() +
  xlab("Residuals")

#Constant Variability
ggplot(data = m_heart, aes(sample = .resid)) +
  stat_qq()
```
```{r}
write.csv(heart3, "C:\\Desktop\\UCLA\\203A\\heartdf.csv", row.names=FALSE)
```

